---
title: "p8105_hw2_sk5855"
author: "Shiyu KANG"
date: "2026-10-02"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
```


```{r problem1 data preparing}
subway_df = 
  readxl::read_xlsx("NYC_Transit_Subway_Entrance_And_Exit_Data.xlsx" , na = c("","NA",".")) |>
  janitor::clean_names()

subway_df_selected = 
  subway_df |>
  select(line:entry, vending, ada) |>
  mutate(
    entry = case_match(entry, "YES" ~ 1, "NO" ~ 0)
  )
```

The dataset contains line, station name and station latitude/longitude, which offers information on the location of stations. Route 1-10 indicates whether and what lines passed this station, 11 lines can pass a single station at most, which is Time Square station. 
For data cleaning, I first uniform the "NA" display. Then uses clean_names to lowercase all variables and add "_" between words. Then select specified variables, using case_match to alter the variable type of entry. The row of the resulting dataset is `r nrow(subway_df_selected)`, the column is `r ncol(subway_df_selected)`. They are untidy. 

```{r problem1 questions}
subway_df_selected = 
  subway_df_selected |>
     mutate(
       line_and_station_name = paste0(line,station_name)
  )

subway_df_selected |> 
  distinct(line_and_station_name) |>
  nrow() 

subway_df_selected |>
  filter(ada == TRUE) |>
  distinct(line_and_station_name) |>
  nrow() 
  
nrow(filter(subway_df_selected,vending == "NO"))/nrow(subway_df_selected)

subway_df_reformat = 
  subway_df_selected |>
  pivot_longer (
    cols = route1:route11,
    names_to = "Route_number",
    names_prefix = "route",
    names_transform = list(Route_number = as.integer),
    values_to = "Route_name",
    values_transform = list(Route_name = as.character),
    values_drop_na = TRUE
  )


  
```

