---
title: "p8105_hw2_sk5855"
author: "Shiyu KANG"
date: "2026-10-02"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
```


```{r problem1 data preparing}
subway_df = 
  readxl::read_xlsx("NYC_Transit_Subway_Entrance_And_Exit_Data.xlsx" , na = c("","NA",".")) |>
  janitor::clean_names()

subway_df_selected = 
  subway_df |>
  select(line:entry, vending, ada) |>
  mutate(
    entry = case_match(entry, "YES" ~ 1, "NO" ~ 0)
  )
```

The dataset contains line, station name and station latitude/longitude, which offers information on the location of stations. Route 1-10 indicates whether and what lines passed this station, 11 lines can pass a single station at most, which is Time Square station. 
For data cleaning, I first uniform the "NA" display. Then uses clean_names to lowercase all variables and add "_" between words. Then select specified variables, using case_match to alter the variable type of entry. The row of the resulting dataset is `r nrow(subway_df_selected)`, the column is `r ncol(subway_df_selected)`. They are untidy. 

```{r problem1 questions}
subway_df_selected = 
  subway_df_selected |>
     mutate(
       line_and_station_name = paste0(line,station_name)
  )

subway_df_selected |> 
  distinct(line_and_station_name) |>
  nrow() 

subway_df_selected |>
  filter(ada == TRUE) |>
  distinct(line_and_station_name) |>
  nrow() 
  
nrow(filter(subway_df_selected,vending == "NO"))/nrow(subway_df_selected)

subway_df_reformat = 
  subway_df_selected |>
  pivot_longer (
    cols = route1:route11,
    names_to = "Route_number",
    names_prefix = "route",
    names_transform = list(Route_number = as.integer),
    values_to = "Route_name",
    values_transform = list(Route_name = as.character),
    values_drop_na = TRUE
  )

subway_df_reformat |>
  filter(Route_name == "A") |>
  distinct(line_and_station_name) |>
  nrow()

subway_df_reformat |>
  filter(Route_name == "A", ada == TRUE) |>
  distinct(line_and_station_name) |>
  nrow()
```

```{r problem 2}
trash_df = 
  readxl::read_xlsx(
    "202509 Trash Wheel Collection Data.xlsx",
    sheet = "Mr. Trash Wheel",
    range = "A2:N709") |>
  janitor::clean_names() |>
  mutate(
    sports_balls = as.integer(round(sports_balls))
  )

professor_df = 
  readxl::read_xlsx(
    "202509 Trash Wheel Collection Data.xlsx",
    sheet = "Professor Trash Wheel",
    range = "A2:M134") |>
  janitor::clean_names() 

gwynns_df = 
  readxl::read_xlsx(
    "202509 Trash Wheel Collection Data.xlsx",
    sheet = "Gwynns Falls Trash Wheel",
    range = "A2:L351") |>
  janitor::clean_names() 

trash_df = trash_df |>
  mutate(
    trash_wheel = "Mr. Trash Wheel"
  )

professor_df = professor_df |>
  mutate(
    trash_wheel = "Professor Trash Wheel"
  )

gwynns_df = gwynns_df |>
  mutate(
    trash_wheel = "Gwynns Falls Trash Wheel"
  )

trash_df = trash_df |>
  mutate(year = as.integer(year))

professor_df = professor_df |>
  mutate(year = as.integer(year))

gwynns_df = gwynns_df |>
  mutate(year = as.integer(year))

combined_df = 
  bind_rows(trash_df, professor_df, gwynns_df)

total_weight = 
  combined_df |>
  filter(trash_wheel == "Professor Trash Wheel") |>
  summarise(total_weight_tons = sum(weight_tons, na.rm = TRUE))

gwynnda_june_butts =
  combined_df |>
  filter(
    trash_wheel == "Gwynns Falls Trash Wheel",
    year == 2022,
    month == "June"
  ) |>
  summarise(
    total_cigarette_butts = sum(cigarette_butts, na.rm = TRUE) 
  ) |> 
      pull(total_cigarette_butts)
```

The combined dataset contains `r nrow(combined_df)` observations and `r ncol(combined_df)` variables from Mr. Trash Wheel, Professor Trash Wheel, and Gwynnda. Each observation represents a dumpster collection, with the `trash_wheel` variable identifying its source. Key variables include `dumpster`, the dumpster number; `date`, the collection date; `weight_tons`, the weight of trash collected in tons; and `volume_cubic_yards`, its volume in cubic yards. The dataset also records counts of specific types of litter, such as `plastic_bottles`, `plastic_bags`, and `sports_balls`. Sports ball counts were rounded to the nearest integer and stored as integers. Variables unavailable for a particular trash wheel are represented by `NA` in the combined dataset.
Professor Trash Wheel collected a total of `r total_weight` tons of trash. Gwynnda collected `r format(gwynnda_june_butts, scientific = FALSE)` cigarette butts in June 2022.
